A quantitative framework for investigating the reliability of network construction

### 2018-01-29 manuscript/CI_interactions.tex
Fixed a few minor spelling mistakes

    In this study, we develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analystical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions [[double-check that we do this]]. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

  Second, where the number of samples $n$ is very low (some pairs of species may never have been documenting as co-occurring), there will be considerable uncertainty around our estimate of $\lambda$. In Fig.~\ref{upper_limits} and Box~\ref{box2}, we derive the Clopper-Pearson interval to explore how the estimate of $\lambda$ varies with sample size. At a small sample size, the 95\% confidence interval spans all values of $\lambda$. To establish that species are not interacting with any acceptable certainty requires tens of observations of the two species co-occurring but not interacting. As most data sets will lack such extensive sampling across all species pairs, we can use a Bayesian approach to supplement what data we do have with other sources of information.

@@

          If our focal network describes a system similar to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. Similar to the procedure for degree distribution, the distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. For distributions of either degrees or interaction probabilities, the procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

  \begin{figure}[h!]
    \caption{A simple example will illustrate the problem of imperfect detection of interactions (the full reasoning is provided below). Assume that we want to infer the probability of an interaction between two species, $i$ and $j$. Now assume that in reality, interaction between $i$ and $j$ is completely impossible (i.e. the true $\lambda=0$), but the observer does not know this and seeks to estimate this interaction probability ($\lambda$). The number of observed interactions will follow a binomial distribution, with number of interactions $k$ and number of observations $n$. Using this distribution, we can compute the credible interval of the estimated probability $\lambda$. Assuming no added detection error in observing the incidence of the interaction, a single observation of species co-occurrence will reveal very little regarding the probability of the interaction as the credible interval for a pair of species with one observation essentially spans from 0 to 1. Only with thirty observations will the uppoer limit of the credible interval be lowered to 0.1. Thus, adding more observations is certainly useful in controlling uncertainty, but the number of observations nonetheless needs to be very high. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for $\lambda$ when $k=0$ ($i$ and $j$ have not been observed interacting) for a variety of $n$ (observed co-occurences of $i$ and $j$). A threshold interaction probability of 0.1 is indicated by the dotted red line.    }
    \label{upper_limits}
    \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
  \end{figure}

>>>>

    In this study, we develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions [[double-check that we do this]]. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

  Second, where the number of samples $n$ is very low (some pairs of species may never have been documented as co-occurring), there will be considerable uncertainty around our estimate of $\lambda$. In Fig.~\ref{upper_limits} and Box~\ref{box2}, we derive the Clopper-Pearson interval to explore how the estimate of $\lambda$ varies with sample size. At a small sample size, the 95\% confidence interval spans all values of $\lambda$. To establish that species are not interacting with any acceptable certainty requires tens of observations of the two species co-occurring but not interacting. As most data sets will lack such extensive sampling across all species pairs, we can use a Bayesian approach to supplement what data we do have with other sources of information.

@@

          If our focal network describes a system similar to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_{ij}$ depends on the degrees of species $i$ and $j$. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_{ij}$=$\Delta_i\times\Delta_j$. Similar to the procedure for degree distribution, the distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. For distributions of either degrees or interaction probabilities, the procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

  \begin{figure}[h!]
    \caption{A simple example will illustrate the problem of imperfect detection of interactions (the full reasoning is provided below). Assume that we want to infer the probability of an interaction between two species, $i$ and $j$. Now assume that in reality, interaction between $i$ and $j$ is completely impossible (i.e. the true $\lambda=0$), but the observer does not know this and seeks to estimate this interaction probability ($\lambda$). The number of observed interactions will follow a binomial distribution, with number of interactions $k$ and number of observations $n$. Using this distribution, we can compute the credible interval of the estimated probability $\lambda$. Assuming no added detection error in observing the incidence of the interaction, a single observation of species co-occurrence will reveal very little regarding the probability of the interaction as the credible interval for a pair of species with one observation essentially spans from 0 to 1. Only with thirty observations will the upper limit of the credible interval be lowered to 0.1. Thus, adding more observations is certainly useful in controlling uncertainty, but the number of observations nonetheless needs to be very high. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for $\lambda$ when $k=0$ ($i$ and $j$ have not been observed interacting) for a variety of $n$ (observed co-occurences of $i$ and $j$). A threshold interaction probability of 0.1 is indicated by the dotted red line.    }
    \label{upper_limits}
    \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
  \end{figure}
